|

### 2024-10-29 output/seminar-slideshow.qmd
Correct IFA probability function formula
Mkes several style updates in plots

library(tibble)
library(dplyr)
library(knitr)
library(lavaan)
library(semPlot)
library(extrafont)
library(tidyr)
library(ggplot2)
library(units)
library(ggtext)
library(ggridges)
library(qgraph)
library(purrr)
library(plotly)
library(metR)
library(scales)
```

@@

GRAPH_FONT   <- "Arial"    # Replace default font to comply with UNED identity
FONT_SIZE    <- 24         # Replace default font size for slideshow
AXIS_LAB_POS <-   .5       # Replace default axis label position for slideshow
LINE_WIDTH   <-   .75      # Replace default line width for slideshow
AXIS_COLOR   <- "darkgray" # Replace default grid line color for slideshow
VECTOR_WIDTH <- 2.5        # Replace default vector width for slideshow
```

```{r graphical-output-conf}
#| cache: false
theme_set( # `ggplot` output configuration
  theme_classic(
    base_size      = FONT_SIZE,
    base_family    = GRAPH_FONT,
    base_line_size = LINE_WIDTH
  ) %+replace%
    theme(
      axis.title.x     = element_text(hjust = AXIS_LAB_POS),
      axis.title.y     = element_text(vjust = AXIS_LAB_POS, angle = 0),
      panel.grid       = element_line(color = AXIS_COLOR),
      panel.grid.minor = element_line(linetype = "17", color = "grey70")
    )
)
```

@@

simple_item_loadings_wide |>
  ggplot(mapping = aes(F1, F2, color = main_factor, label = item)) +
  geom_point(size = I(4))                                          +
  geom_text(
    nudge_x =  .1,
    nudge_y = -.08,
    color   = I("black"),
    size    = I(5)
  )                                                                +
  scale_x_continuous(
    limits = axis_lims,
    breaks = 0,
    labels = NULL,
    name   = "F2",
  )                                                                +
  scale_y_continuous(
    limits = axis_lims,
    breaks = 0,
    labels = NULL,
    name   = "F1"
  )                                                                +
  scale_color_manual(values = factor_colors, guide = NULL)         +
  coord_fixed(expand = FALSE, clip = "on")                         +
  theme(
    axis.line          = element_blank(),
    axis.ticks         = element_blank(),
    axis.title.x       = element_text(hjust = axis_pos),
    axis.title.y       = element_text(vjust = axis_pos),
    panel.grid.major   = element_line(
      color     = "black",
      linewidth = LINE_WIDTH
    ),
    panel.grid.major.x = element_line(color = PALETTE_RASPBERRY["light"]),
    panel.grid.major.y = element_line(color = PALETTE_BLUE["light"])
  )
```
:::
:::::

@@

simple_item_loadings_rot_plot <- simple_item_loadings_rot |>
  group_by(item)                                          |>
  pivot_wider(names_from = factor, values_from = loading) |>
  ggplot(mapping = aes(F1, F2, color = main_factor, label = item)) +
  geom_point(size = I(4))                                          +
  geom_text(
    nudge_x = -.12,
    # nudge_y = -.08,
    color   = I("black"),
    size    = I(5)
  )                                                                +
  scale_x_continuous(
    limits = axis_lims,
    breaks = 0,
    labels = NULL,
    name   = "",
  )                                                                +
  scale_y_continuous(
    limits = axis_lims,
    breaks = 0,
    labels = NULL,
    name   = ""
  )                                                                +
  scale_color_manual(values = factor_colors, guide = NULL)         +
  coord_fixed(expand = FALSE, clip = "on")                         +
  theme(
    axis.line          = element_blank(),
    axis.ticks         = element_blank(),
    axis.title.x       = element_text(hjust = axis_pos),
    axis.title.y       = element_text(vjust = axis_pos),
    panel.grid.major   = element_line(
      color     = "black",
      linewidth = LINE_WIDTH
    )
  )

simple_item_loadings_rot_plot
```
:::

@@

simple_item_loadings_rot_plot +
  geom_abline(
    data    = rotated_axes,
    mapping = aes(slope = slope, intercept = 0, color = factor)
  )
```
:::

@@

complex_item_loadings_wide |>
  ggplot(mapping = aes(F1, F2, color = main_factor, label = item)) +
  geom_segment(
    data    = item3_spikes,
    mapping = aes(F1_start, F2_start, xend = F1_end, yend = F2_end),
    inherit.aes = FALSE,
    linewidth = LINE_WIDTH,
    linetype  = '43',
    color = spike_colors
  )                                                                +
  geom_point(size = I(4))                                          +
  geom_text(
    nudge_x =  .1,
    nudge_y = -.08,
    color   = I("black"),
    size    = I(5)
  )                                                                +
  scale_x_continuous(
    limits = axis_lims,
    breaks = 0,
    labels = NULL,
    name   = "F2",
  )                                                                +
  scale_y_continuous(
    limits = axis_lims,
    breaks = 0,
    labels = NULL,
    name   = "F1"
  )                                                                +
  scale_color_manual(values = factor_colors, guide = NULL)         +
  coord_fixed(expand = FALSE, clip = "on")                         +
  theme(
    axis.line          = element_blank(),
    axis.ticks         = element_blank(),
    axis.title.x       = element_text(hjust = axis_pos),
    axis.title.y       = element_text(vjust = axis_pos),
    panel.grid.major   = element_line(
      color     = "black",
      linewidth = LINE_WIDTH
    ),
    panel.grid.major.x = element_line(color = PALETTE_RASPBERRY["light"]),
    panel.grid.major.y = element_line(color = PALETTE_BLUE["light"])
  )
```
:::
:::::

@@

method_item_loadings                                      |>
  pivot_wider(names_from = factor, values_from = loading) |>
  ggplot(mapping = aes(F1, M1, label = item))              +
  geom_point(size = I(4), color = F1_COLOR)                +
  geom_text(
    # nudge_x =  .1,
    nudge_y = -.12,
    color   = I("black"),
    size    = I(5)
  )                                                        +
  scale_x_continuous(
    limits = axis_lims,
    breaks = 0,
    labels = NULL,
    name   = "M1",
  )                                                        +
  scale_y_continuous(
    limits = axis_lims,
    breaks = 0,
    labels = NULL,
    name   = "F1"
  )                                                        +
  scale_color_manual(values = factor_colors, guide = NULL) +
  coord_fixed(expand = FALSE, clip = "on")                 +
  theme(
    axis.line          = element_blank(),
    axis.ticks         = element_blank(),
    axis.title.x       = element_text(hjust = axis_pos),
    axis.title.y       = element_text(vjust = axis_pos),
    panel.grid.major   = element_line(
      color     = "black",
      linewidth = LINE_WIDTH
    ),
    panel.grid.major.x = element_line(color = PALETTE_RASPBERRY["light"]),
    panel.grid.major.y = element_line(color = PALETTE_BLUE["light"])
  )
```
:::
:::::

@@

::: {.column .fragment fragment-index="2"}
```{r fa_example_plot}
#| cache: false
AXIS_MINOR <- c(-1, 1, 2)

@@

fa_example_wide |>
  ggplot(mapping = aes(F1, F2, label = object))            +
  geom_segment(
    data        = fa_example_spikes,
    mapping     = aes(
      F1_start, F2_start,
      xend  = F1_end, yend = F2_end
    ),
    inherit.aes = FALSE,
    linewidth   = LINE_WIDTH,
    linetype    = '43',
    color       = spike_colors |> rep(each = 2)
  )                                                        +
  geom_point(size = I(4))                                  +
  geom_richtext(
    nudge_x = .15,
    nudge_y = .10,
    color   = I("black"),
    size    = I(5),
    lineheight = 1.2,
    fill = NA,
    label.color = NA
  )                                                        +
  scale_x_continuous(
    limits       = axis_lims,
    breaks       = 0,
    labels       = NULL,
    name         = "F~2~",
    minor_breaks = AXIS_MINOR
  )                                                        +
  scale_y_continuous(
    limits       = axis_lims,
    breaks       = 0,
    labels       = NULL,
    name         = "F~1~",
    minor_breaks = AXIS_MINOR
  )                                                        +
  scale_color_manual(values = factor_colors, guide = NULL) +
  coord_fixed(expand = FALSE, clip = "on")                 +
  theme(
    axis.line          = element_blank(),
    axis.ticks         = element_blank(),
    axis.title.x       = element_markdown(hjust = axis_pos),
    axis.title.y       = element_markdown(vjust = axis_pos),
    panel.grid.major   = element_line(
      color     = "black",
      linewidth = LINE_WIDTH
    ),
    panel.grid.major.x = element_line(color = PALETTE_RASPBERRY["light"]),
    panel.grid.major.y = element_line(color = PALETTE_BLUE["light"])
  )
```
:::
:::::::

@@

::::::::: columns
::::: {.column width="40%"}
$$
  y_i^* = \tau_i + \mathbf{\lambda}_i^T \mathbf{F} + U_i
$$

@@

::: {.fragment fragment-index="2"}
$$
  P(Y_i = 1) = \int_{-\infty}^0 \mathcal{N}(y_i^*) \delta y_i^* \\
             = \Phi_N \left(\frac{\tau_i + \mathbf{\lambda}_i^T \mathbf{F}}{\sqrt{var(U_i)}}\right)
$$
:::
:::::

::::: {.column width="60%"}
::: {.fragment fragment-index="1"}
```{r ifa-density}
#| fig-asp:   0.5
#| fig-width: 8

@@

# Plot: Upper Panel - Linear regression with Gaussian error distributions
latent_response |>
  ggplot(
    aes(x = response, y = point, fill = after_stat(x) > 0, group = point)
  )                            +
  geom_density_ridges_gradient(
    scale          =  .2,
    rel_min_height = 0.001,
    color          = PALETTE_BLUE["normal"]
  )                              +
  labs(x = "y~i~^*^", y = NULL)  +
  scale_y_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    labels       = NULL,
    limits       = range(ERROR_POINTS) * 1.12,
    expand       = expansion()
  )           +
  scale_x_continuous(breaks = 0) +
  scale_fill_manual(
    values = c(
      `FALSE` = unname(PALETTE_BLUE["white"]),
      `TRUE`  = unname(PALETTE_BLUE["normal"])
    )
  )                              +
  theme(
    axis.ticks.x = element_blank(),
    axis.title.y = element_markdown(angle = 90),
    legend.position = "none",
    panel.grid.major = element_line(linetype = "17")
  )                              +
  coord_flip()
```
:::

@@

# Plot: Lower Panel - Probit curve
probit_curve_data |>
  ggplot(mapping = aes(x = score, y = prob))                        +
  geom_line(color = PALETTE_BLUE["normal"], linewidth = LINE_WIDTH) +
  labs(x = "E(*y~i~^*^*)", y = "P(Y~i~ = 1)")                       +
  scale_y_continuous(
    breaks       = 0:1,
    minor_breaks = NULL,
    limits       = 0:1,
    expand       = expansion(mult = c(0, .05))
  )  +
  scale_x_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    limits = range(ERROR_POINTS) * 1.12,
    expand = expansion()
  ) +
  theme(
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(angle = 90),
    panel.grid.major = element_line(linetype = "17"),
  )
```
:::
:::::
:::::::::

@@

logit_curve_data <- tibble(
  latent_trait = PROB_AXIS,
  prob         = ITEMS |>
    select(-item) |>
    pmap(irf, theta = latent_trait) |>
    bind_cols()
) |>
  unnest(
    prob,
    names_sep = '_'
  ) |>
  rename_with(stringr::str_remove, pattern = '\\.{3}')

@@

logit_curve_data_long |>
  ggplot(
    mapping = aes(
      latent_trait, prob,
      color = item,
      group = after_scale(color)
    )
  )                                        +
  geom_line(linewidth = LINE_WIDTH)        +
  labs(x = "*&theta;*", y = "P(Y~i~ = 1)") +
  scale_y_continuous(
    breaks       = 0:1,
    minor_breaks = NULL,
    limits       = 0:1,
    expand       = expansion(mult = c(0, .05))
  )                                        +
  scale_x_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    limits = range(ERROR_POINTS) * 1.12,
    expand = expansion()
  )                                        +
  scale_color_manual(values = ITEMS_PALETTE, guide = NULL) +
  theme(
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(angle = 90),
    panel.grid.major = element_line(linetype = "17"),
  )
```

@@

::: column
```{r mirt-model-irs-contour}
contour_plot <- mirt_item_probs |>
  ggplot(mapping = aes(trait_1, trait_2, z = prob, label = after_stat(level))) +
  geom_contour2(
    color = F1_COLOR,
    linewidth = LINE_WIDTH,
    breaks = c(.05, .1, .2, .3, .5, .7, .8, .9, .95),
    label.placer = label_placer_fraction(frac = .18, rot_adjuster = 0)
  ) +
  scale_x_continuous(
    minor_breaks = NULL,
    limits = range(PROB_AXIS_2D),
    expand = expansion()
  ) +
  scale_y_continuous(
    minor_breaks = NULL,
    limits = range(PROB_AXIS_2D),
    expand = expansion()
  ) +
  coord_fixed() +
  labs(x = "*&theta;*~1~", y = "*&theta;*~2~") +
  theme(
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(angle = 90),
    panel.grid.major = element_line(linetype = "17"),
  )

@@

:::: column
\begin{matrix}
  \mathbf{\theta} =
    \begin{bmatrix}
       1 \\
      -1
    \end{bmatrix}

@@

mirt_item_coords <- mirt_item_params |>
  compute_mirt_coords(D, MDISC, starts_with("cos"), original_coords = FALSE)

mirt_item_spikes <- mirt_item_coords |>
  select(item, trait_1 = origin_transf_1, trait_2 = origin_transf_2) |>
  bind_rows(tibble(item = 1, trait_1 = 0, trait_2 = 0)) |>
  add_column(type = c("end", "start")) |>
  pivot_wider(names_from = type, values_from = starts_with("trait"))

contour_plot +
    geom_segment(
    data    = mirt_item_spikes,
    mapping = aes(
      trait_1_start, trait_2_start,
      xend = trait_1_end,
      yend = trait_2_end
    ),
    inherit.aes = FALSE,
    linewidth = LINE_WIDTH,
    linetype  = '43',
    color = PALETTE_BLUE["light"]
  ) +
  geom_point(
    data = mirt_item_coords,
    mapping = aes(origin_transf_1, origin_transf_2),
    inherit.aes = FALSE,
    size = I(4)
  ) +
  scale_x_continuous(
    breaks = -1:1,
    minor_breaks = NULL,
    limits = c(-1.5, 1.5)
  ) +
  scale_y_continuous(
    breaks = -1:1,
    minor_breaks = NULL,
    limits = c(-1.5, 1.5)
  ) +
  annotate(
    geom = "text",
    x = 0.1, y = 0.2,
    label = "MID",
    size = 6,
    angle = mirt_item_params |> pull(deg_1)
  )
```
:::

@@

> the discriminating power of an item indicates how quickly the transition takes place from low probability to high probability of a correct response.
> \[...\] A highly discriminating item divides the regions clearly-having a narrow region of ambiguity, that is, a region where the probabilities are intermediate in magnitude.

::::: {.columns .fragment}
::: column
<br>

@@

::: column
```{r mdisc-item}
#| fig-width:  4
#| fig-height: 4
mirt_item_coords |>
  arrange(desc(item))   |> # To plot them in reverse order
  ggplot(
    aes(
      0, 0,
      xend = end_transf_1 - origin_transf_1,
      yend = end_transf_2 - origin_transf_2,
      color = factor(item), fill = factor(item)
    ),
  )                                                  +
  geom_vline(xintercept = 0, linewidth = LINE_WIDTH) +
  geom_hline(yintercept = 0, linewidth = LINE_WIDTH) +
  geom_segment(
    arrow     = arrow(angle = 20, length = unit(10, "points"), type = "closed"),
    linejoin  = "mitre",
    linewidth = VECTOR_WIDTH
  ) +
  scale_x_continuous(
    limits       = c(-1, 2),
    breaks       = -1:2,
    minor_breaks = NULL,
    name         = NULL,
    oob          = oob_keep
  ) +
  scale_y_continuous(
    limits       = c(-1, 2),
    breaks       = -1:2,
    minor_breaks = NULL,
    name         = NULL,
    oob          = oob_keep
  ) +
  scale_color_manual(values = unname(PALETTE_BLUE["normal"]), guide = NULL) +
  coord_fixed(expand = FALSE, clip = "on")           +
  theme(
    axis.line        = element_blank(),
    panel.grid.major = element_line(
      linewidth = LINE_WIDTH,
      linetype  = "17"
    )
  )
```
:::
:::::

>>>>

library(tibble)
library(dplyr)
library(knitr)
library(lavaan)
library(semPlot)
library(extrafont)
library(tidyr)
library(ggplot2)
library(units)
library(ggtext)
library(ggridges)
library(qgraph)
library(stringr)
library(purrr)
library(plotly)
library(metR)
library(scales)
```

@@

GRAPH_FONT   <- "Arial"  # Replace default font to comply with UNED identity
FONT_SIZE    <- 24       # Replace default font size for slideshow
AXIS_LAB_POS <-  0.5     # Replace default axis label position for slideshow
LINE_WIDTH   <-  0.75    # Replace default line width for slideshow
GRID_WIDTH   <-  0.5     # Width for grid and axis lines
AXIS_COLOR   <- "gray70" # Replace default grid line color for slideshow
VECTOR_WIDTH <- 1.5      # Replace default vector width for slideshow
```

```{r graphical-output-conf}
#| cache: false
theme_set( # `ggplot` output configuration
  theme_classic(
    base_size      = FONT_SIZE,
    base_family    = GRAPH_FONT,
    base_line_size = LINE_WIDTH
  ) %+replace%
    theme(
      axis.title.x       = element_text(hjust = AXIS_LAB_POS),
      axis.title.y       = element_text(vjust = AXIS_LAB_POS, angle = 0),
      axis.title.y.right = element_text(vjust = AXIS_LAB_POS, angle = 0),
      panel.grid         = element_line(color = AXIS_COLOR),
      panel.grid.minor   = element_line(linetype = "17", color = AXIS_COLOR)
    )
)
```

@@

simple_item_loadings_wide |>
  ggplot(mapping = aes(F1, F2, color = main_factor, label = item)) +
  geom_point(size = I(4))                                          +
  geom_text(
    nudge_x =  .1,
    nudge_y = -.08,
    color   = I("black"),
    size    = I(5)
  )                                                                +
  scale_x_continuous(
    limits   = axis_lims,
    breaks   = 0,
    labels   = NULL,
    name     = "F2",
    position = "top"
  )                                                                +
  scale_y_continuous(
    limits   = axis_lims,
    breaks   = 0,
    labels   = NULL,
    name     = "F1",
    position = "right"
  )                                                                +
  scale_color_manual(values = factor_colors, guide = NULL)         +
  coord_fixed(expand = FALSE, clip = "on")                         +
  theme(
    axis.line          = element_blank(),
    axis.ticks         = element_blank(),
    axis.title.x       = element_text(hjust = axis_pos),
    axis.title.y.right = element_text(vjust = axis_pos),
    panel.grid.major   = element_line(
      color     = "black",
      linewidth = GRID_WIDTH
    ),
    panel.grid.major.x = element_line(color = PALETTE_RASPBERRY["light"]),
    panel.grid.major.y = element_line(color = PALETTE_BLUE["light"])
  )
```
:::
:::::

@@

simple_item_loadings_rot_plot <- simple_item_loadings_rot |>
  group_by(item)                                          |>
  pivot_wider(names_from = factor, values_from = loading) |>
  ggplot(mapping = aes(F1, F2, color = main_factor, label = item)) +
  geom_text(
    nudge_x = -.12,
    color   = I("black"),
    size    = I(5)
  )                                                                +
  scale_x_continuous(
    limits = axis_lims,
    breaks = 0,
    labels = NULL,
    name   = "",
  )                                                                +
  scale_y_continuous(
    limits = axis_lims,
    breaks = 0,
    labels = NULL,
    name   = ""
  )                                                                +
  scale_color_manual(values = factor_colors, guide = NULL)         +
  coord_fixed(expand = FALSE, clip = "on")                         +
  theme(
    axis.line          = element_blank(),
    axis.ticks         = element_blank(),
    panel.grid.major   = element_line(
      color     = "black",
      linewidth = GRID_WIDTH
    )
  )

simple_item_loadings_rot_plot + geom_point(size = I(4))
```
:::

@@

simple_item_loadings_rot_plot +
  geom_abline(
    data    = rotated_axes,
    mapping = aes(slope = slope, intercept = 0),
    color   = c(PALETTE_BLUE["light"], PALETTE_RASPBERRY["light"]),
    linewidth = LINE_WIDTH
  ) +
  geom_point(size = I(4)) +
  annotate(
    "text",
    x = -1.35, y = c(-.62, .78),
    label = rotated_axes |> pull(factor),
    size  = 9
  )
```
:::

@@

complex_item_loadings_wide |>
  ggplot(mapping = aes(F1, F2, color = main_factor, label = item)) +
  geom_segment(
    data    = item3_spikes,
    mapping = aes(F1_start, F2_start, xend = F1_end, yend = F2_end),
    inherit.aes = FALSE,
    linewidth = LINE_WIDTH,
    linetype  = '43',
    color = spike_colors
  )                                                                +
  geom_point(size = I(4))                                          +
  geom_text(
    nudge_x =  .1,
    nudge_y = -.08,
    color   = I("black"),
    size    = I(5)
  )                                                                +
  scale_x_continuous(
    limits = axis_lims,
    breaks = 0,
    labels = NULL,
    name   = "F2",
    position = "top"
  )                                                                +
  scale_y_continuous(
    limits = axis_lims,
    breaks = 0,
    labels = NULL,
    name   = "F1",
    position = "right"
  )                                                                +
  scale_color_manual(values = factor_colors, guide = NULL)         +
  coord_fixed(expand = FALSE, clip = "on")                         +
  theme(
    axis.line          = element_blank(),
    axis.ticks         = element_blank(),
    axis.title.x       = element_text(hjust = axis_pos),
    axis.title.y.right = element_text(vjust = axis_pos),
    panel.grid.major   = element_line(
      color     = "black",
      linewidth = GRID_WIDTH
    ),
    panel.grid.major.x = element_line(color = PALETTE_RASPBERRY["light"]),
    panel.grid.major.y = element_line(color = PALETTE_BLUE["light"])
  )
```
:::
:::::

@@

method_item_loadings                                      |>
  pivot_wider(names_from = factor, values_from = loading) |>
  ggplot(mapping = aes(F1, M1, label = item))              +
  geom_point(size = I(4), color = F1_COLOR)                +
  geom_text(
    nudge_y = -.12,
    color   = I("black"),
    size    = I(5)
  )                                                        +
  scale_x_continuous(
    limits = axis_lims,
    breaks = 0,
    labels = NULL,
    name   = "M1",
    position = "top"
  )                                                        +
  scale_y_continuous(
    limits = axis_lims,
    breaks = 0,
    labels = NULL,
    name   = "F1",
    position = "right"
  )                                                        +
  scale_color_manual(values = factor_colors, guide = NULL) +
  coord_fixed(expand = FALSE, clip = "on")                 +
  theme(
    axis.line          = element_blank(),
    axis.ticks         = element_blank(),
    axis.title.x       = element_text(hjust = axis_pos),
    axis.title.y.right = element_text(vjust = axis_pos),
    panel.grid.major   = element_line(
      color     = "black",
      linewidth = GRID_WIDTH
    ),
    panel.grid.major.x = element_line(color = PALETTE_RASPBERRY["light"]),
    panel.grid.major.y = element_line(color = PALETTE_BLUE["light"])
  )
```
:::
:::::

@@

::: {.column .fragment fragment-index="2"}
```{r fa_example_plot}
#| cache: false
AXIS_GRID <- -1:1

@@

fa_example_wide |>
  ggplot(mapping = aes(F1, F2, label = object))             +
  geom_hline(
    yintercept = 0,
    color      = PALETTE_BLUE["light"],
    linewidth  = GRID_WIDTH
  )                                                         +
  geom_vline(
    xintercept = 0,
    color      = PALETTE_RASPBERRY["light"],
    linewidth  = GRID_WIDTH
  )                                                         +
  geom_segment(
    data        = fa_example_spikes,
    mapping     = aes(
      F1_start, F2_start,
      xend  = F1_end, yend = F2_end
    ),
    inherit.aes = FALSE,
    linewidth   = LINE_WIDTH,
    linetype    = '43',
    color       = spike_colors |> rep(each = 2)
  )                                                        +
  geom_point(size = I(4))                                  +
  geom_richtext(
    nudge_x = .15,
    nudge_y = .10,
    color   = I("black"),
    size    = I(5),
    lineheight = 1.2,
    fill = NA,
    label.color = NA
  )                                                        +
  scale_x_continuous(
    limits       = axis_lims,
    breaks       = AXIS_GRID,
    name         = NULL,
    minor_breaks = NULL,
    sec.axis     = dup_axis(name = "F~2~", labels = NULL)
  )                                                        +
  scale_y_continuous(
    limits       = axis_lims,
    breaks       = AXIS_GRID,
    name         = NULL,
    minor_breaks = NULL,
    sec.axis     = dup_axis(name = "F~1~", labels = NULL)
  )                                                        +
  scale_color_manual(values = factor_colors, guide = NULL) +
  coord_fixed(expand = FALSE, clip = "on")                 +
  theme(
    axis.line          = element_blank(),
    axis.ticks         = element_blank(),
    axis.title.x       = element_markdown(hjust = axis_pos),
    axis.title.y       = element_markdown(vjust = axis_pos),
    axis.title.y.right = element_text(vjust = axis_pos),
    panel.grid.major   = element_line(
      linewidth = GRID_WIDTH,
      linetype  = "17"
    )
  )
```
:::
:::::::

@@

::::::::: columns
::::: {.column width="45%"}
$$
  y_i^* = \tau_i + \mathbf{\lambda}_i^T \mathbf{F} + U_i
$$

@@

::: {.fragment fragment-index="2"}
$$
  \begin{align}
    P(Y_i = 1) &= \int_0^{-\infty} \mathcal{N}(y_i^*) \delta y_i^* \\
               &= \Phi_N \left(
                 \frac{ \tau_i + \mathbf{\lambda}_i^T \mathbf{F} }
                   { \sqrt{var(U_i)} }
               \right)
  \end{align}
$$
:::
:::::

::::: {.column width="55%"}
::: {.fragment fragment-index="1"}
```{r ifa-density}
#| fig-asp:   0.5
#| fig-width: 8

@@

# Plot: Upper Panel - Linear regression with Gaussian error distributions
latent_response |>
  ggplot(
    aes(x = response, y = point, fill = after_stat(x) > 0, group = point)
  )                            +
  geom_density_ridges_gradient(
    scale          =  .2,
    rel_min_height = 0.001,
    color          = PALETTE_BLUE["normal"]
  )                              +
  labs(x = "y~i~^*^", y = NULL)  +
  scale_y_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    labels       = NULL,
    limits       = range(ERROR_POINTS) * 1.12,
    expand       = expansion()
  )           +
  scale_x_continuous(breaks = 0) +
  scale_fill_manual(
    values = c(
      `FALSE` = unname(PALETTE_BLUE["white"]),
      `TRUE`  = unname(PALETTE_BLUE["normal"])
    )
  )                              +
  theme(
    axis.ticks.x = element_blank(),
    axis.title.y = element_markdown(angle = 90),
    legend.position = "none",
    panel.grid.major = element_line(linetype = "17", linewidth = GRID_WIDTH)
  )                              +
  coord_flip()
```
:::

@@

# Plot: Lower Panel - Probit curve
probit_curve_data |>
  ggplot(mapping = aes(x = score, y = prob))                        +
  geom_line(color = PALETTE_BLUE["normal"], linewidth = LINE_WIDTH) +
  labs(x = "E(*y~i~^*^*)", y = "P(Y~i~ = 1)")                       +
  scale_y_continuous(
    breaks       = 0:1,
    minor_breaks = NULL,
    limits       = 0:1,
    expand       = expansion(mult = c(0, .05)),
    oob = oob_keep
  )  +
  scale_x_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    limits = range(ERROR_POINTS) * 1.12,
    expand = expansion()
  ) +
  theme(
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(angle = 90),
    panel.grid.major = element_line(linetype = "17", linewidth = GRID_WIDTH),
  )
```
:::
:::::
:::::::::

@@

logit_curve_data <- tibble(
  latent_trait = PROB_AXIS,
  prob         = ITEMS |>
    select(-item) |>
    pmap(irf, theta = latent_trait) |>
    bind_cols()
) |>
  unnest(
    prob,
    names_sep = '_'
  ) |>
  rename_with(str_remove, pattern = '\\.{3}')

@@

logit_curve_data_long |>
  ggplot(
    mapping = aes(
      latent_trait, prob,
      color = item,
      group = after_scale(color)
    )
  )                                        +
  geom_line(linewidth = LINE_WIDTH)        +
  labs(x = "*&theta;*", y = "P(Y~i~ = 1)") +
  scale_y_continuous(
    breaks       = 0:1,
    minor_breaks = NULL,
    limits       = 0:1,
    expand       = expansion(mult = c(0, .05))
  )                                        +
  scale_x_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    limits = range(ERROR_POINTS) * 1.12,
    expand = expansion()
  )                                        +
  scale_color_manual(values = ITEMS_PALETTE, guide = NULL) +
  theme(
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(angle = 90),
    panel.grid.major = element_line(linetype = "17", linewidth = GRID_WIDTH),
  )
```

@@

::: column
```{r mirt-model-irs-contour}
contour_plot <- mirt_item_probs |>
  ggplot(mapping = aes(trait_1, trait_2, z = prob, label = after_stat(level))) +
  geom_vline(xintercept = 0, linewidth = GRID_WIDTH) +
  geom_hline(yintercept = 0, linewidth = GRID_WIDTH) +
  geom_contour2(
    color = F1_COLOR,
    linewidth = LINE_WIDTH,
    breaks = c(.05, .1, .2, .3, .5, .7, .8, .9, .95),
    label.placer = label_placer_fraction(frac = .90, rot_adjuster = 0)
  ) +
  scale_x_continuous(
    minor_breaks = NULL,
    limits = range(PROB_AXIS_2D),
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "*&theta;*~2~", labels = NULL)
  ) +
  scale_y_continuous(
    minor_breaks = NULL,
    limits = range(PROB_AXIS_2D),
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "*&theta;*~1~", labels = NULL)
  ) +
  coord_fixed() +
  theme(
    axis.ticks       = element_blank(),
    axis.line        = element_blank(),
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(),
    panel.grid.major = element_line(linetype = "17", linewidth = GRID_WIDTH)
  )

@@

:::: column

\begin{matrix}
  \mathbf{\theta} =
    \begin{bmatrix}
      1 \\
      -1
    \end{bmatrix}

@@

mirt_item_coords <- mirt_item_params |>
  compute_mirt_coords(D, MDISC, starts_with("cos"), original_coords = FALSE) |>
    rename_with(str_remove, pattern = "transf_")

mirt_item_spikes <- mirt_item_coords |>
  select(item, trait_1 = origin_1, trait_2 = origin_2) |>
  bind_rows(tibble(item = 1, trait_1 = 0, trait_2 = 0)) |>
  add_column(type = c("end", "start")) |>
  pivot_wider(names_from = type, values_from = starts_with("trait"))

contour_plot +
  geom_segment(
    data    = mirt_item_spikes,
    mapping = aes(
      trait_1_start, trait_2_start,
      xend = trait_1_end,
      yend = trait_2_end
    ),
    inherit.aes = FALSE,
    linewidth = LINE_WIDTH,
    linetype  = '43',
    color = PALETTE_BLUE["light"]
  ) +
  geom_point(
    data = mirt_item_coords,
    mapping = aes(origin_1, origin_2),
    inherit.aes = FALSE,
    size = I(4)
  ) +
  scale_x_continuous(
    breaks       = -1:1,
    minor_breaks = NULL,
    limits       = c(-1.5, 1.5),
    expand       = expansion(),
    name         = NULL,
    sec.axis     = dup_axis(name = "*&theta;*~2~", labels = NULL)
  ) +
  scale_y_continuous(
    breaks       = -1:1,
    minor_breaks = NULL,
    limits       = c(-1.5, 1.5),
    expand       = expansion(),
    name         = NULL,
    sec.axis     = dup_axis(name = "*&theta;*~1~", labels = NULL)
  ) +
  annotate(
    geom = "text",
    x = 0.18, y = 0.17,
    label = "MID",
    size = 6,
    angle = mirt_item_params |> pull(deg_1)
  )
```
:::

@@

@reckase_discriminating_1991 [p. 362]:

> the discriminating power of an item indicates how quickly the transition takes place from low probability to high probability \[...\].
> A highly discriminating item divides the regions clearly---having a narrow region  \[...\] where the probabilities are intermediate in magnitude.

::::: {.columns .fragment}
::: column
<br>
<br>
<br>

@@

::: column
```{r mdisc-item}
#| fig-width:  4.7
#| fig-height: 4.7
#| cache: false

item_vector <- mirt_item_coords |> mutate(
  across(ends_with("1"), ~ . - origin_1),
  across(ends_with("2"), ~ . - origin_2)
)

item_spikes <- item_vector             |>
  slice(1 |> rep(2))                   |>
  add_column(dim = 'F' |> paste0(1:2)) |>
  mutate(
    origin_1 = if_else(dim == 'F1', end_1, origin_1),
    origin_2 = if_else(dim == 'F2', end_2, origin_2),
  )
